\documentclass[onecolumn, amsmath, amssymb, aps, pra]{revtex4-2}

\usepackage[colorlinks=true,linkcolor=blue,allcolors=blue]{hyperref}

% Verbatim display of the .bib source of each variant
\usepackage{fancyvrb}
\fvset{fontsize=\small, xleftmargin=1em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{How BibTeX entry variants for preprints render under REVTeX}
\date{\today}
\author{Michael Goerz}
\maketitle

This document compares how different ways of writing the \emph{same}
references in a \texttt{.bib} file render in the bibliography of a typical
REVTeX document (\texttt{revtex4-2} with the default \texttt{apsrev4-2}
bibliography style, \texttt{hyperref} loaded). The references are an arXiv
\emph{preprint-only} paper (Wilhelm et al., arXiv:2003.10132), a HAL
preprint-only paper (Turinici, HAL:hal-00640217), and a \emph{published}
paper for which a preprint exists (Koch et al., EPJ Quantum Technol.\ 9, 19
(2022); arXiv:2205.12110). The variants are the forms that
\texttt{bibdeskparser} stores and exports, arXiv's own BibTeX export, and
two instructive anti-patterns. Each variant shows its \texttt{.bib} source,
followed by the citation it produces; compare the numbered entries in the
bibliography at the end. This document is part of the
\texttt{bibdeskparser} documentation; see the ``Preprints'' page there for
the discussion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preprint-only entry (arXiv)}

\subsection{P1: The stored form (\texttt{@unpublished} with
pseudo-journal, eprint fields, DOI, and status
note)~\cite{WilhelmStored}}

How \texttt{bibdeskparser} (and BibDesk) store a preprint-only entry, and
what any full export writes. This is the form to copy verbatim into an
external \texttt{.bib} file: BibTeX ignores the \texttt{journal} field of
an \texttt{@unpublished} entry, so the pseudo-journal (which serves
BibDesk's table view and \texttt{bibdeskparser}'s own rendering) is
invisible here. The \texttt{note} records the publication status
(``preprint only'', ``submitted to Phys.\ Rev.\ A'', ``lecture notes'');
\texttt{bibdeskparser} never fills it in automatically -- a missing note is
the signal to do so by hand.

\begin{Verbatim}
@unpublished{WilhelmStored,
    Author = {Wilhelm, Frank K. and ...},
    Title = {An introduction into optimal control for quantum technologies},
    Journal = {arXiv:2003.10132},
    Eprint = {2003.10132},
    Archiveprefix = {arXiv},
    Primaryclass = {quant-ph},
    Doi = {10.48550/arxiv.2003.10132},
    Note = {lecture notes},
    Year = {2020},
}
\end{Verbatim}

\subsection{P2: The minimal \texttt{--preprint unpublished} export (the
default)~\cite{WilhelmUnpublished}}

What \verb|bibdeskparser export --minimal| produces: as P1, without the
pseudo-journal. For an entry without a stored note, the required
\texttt{note} is synthesized as ``preprint'' (in minimal exports only).

\begin{Verbatim}
@unpublished{WilhelmUnpublished,
    ... same fields as P1, without Journal ...
}
\end{Verbatim}

\subsection{P3: The minimal \texttt{--preprint misc}
export~\cite{WilhelmMisc}}

The same structured fields as a \texttt{@misc} entry (\texttt{@misc} has
no required fields, so a note is kept but never synthesized).

\begin{Verbatim}
@misc{WilhelmMisc,
    ... same fields as P2, as @misc ...
}
\end{Verbatim}

\subsection{P4: The minimal \texttt{--preprint article}
export~\cite{WilhelmArticle}}

The pseudo-journal in the journal position, hyperlinked via the
DOI-resolver address in \texttt{url}; no eprint/doi fields. This is the
form for classic styles (\texttt{plain}, \texttt{unsrt},
\texttt{IEEEtran}, ...), which would silently drop an \texttt{eprint}.

\begin{Verbatim}
@article{WilhelmArticle,
    Author = {Wilhelm, Frank K. and ...},
    Title = {An introduction into optimal control for quantum technologies},
    Journal = {arXiv:2003.10132},
    Url = {https://doi.org/10.48550/arxiv.2003.10132},
    Note = {lecture notes},
    Year = {2020},
}
\end{Verbatim}

\subsection{P5: arXiv's own BibTeX export
(\texttt{@misc})~\cite{WilhelmArxivExport}}

Verbatim what the ``Export BibTeX citation'' button on the arXiv abstract
page produces (\texttt{bibdeskparser import} accepts this form and
normalizes it to P1, deriving the DOI; only the \texttt{note} is left
to fill in by hand).

\begin{Verbatim}
@misc{WilhelmArxivExport,
      title={An introduction into optimal control for quantum technologies},
      author={Frank K. Wilhelm and Susanna Kirchhoff and Shai Machnes
              and Nicolas Wittler and Dominique Sugny},
      year={2020},
      eprint={2003.10132},
      archivePrefix={arXiv},
      primaryClass={quant-ph},
      url={https://arxiv.org/abs/2003.10132},
}
\end{Verbatim}

\subsection{P6: Anti-pattern -- the legacy \texttt{@article} storage
form, copied verbatim~\cite{WilhelmLegacyArticle}}

The same fields as P1, but on an \texttt{@article} (the recognized legacy
storage form). Copied \emph{verbatim} into a REVTeX document instead of
going through \texttt{export}, the journal field is no longer ignored, and
the identifier renders three times.

\begin{Verbatim}
@article{WilhelmLegacyArticle,
    ... same fields as P1, as @article, without Note ...
}
\end{Verbatim}

\subsection{P7: Anti-pattern -- \texttt{@article} with pseudo-journal
and a \texttt{doi} field~\cite{WilhelmArticleDoi}}

Why the \texttt{--preprint article} export (P4) writes the DOI as a
\texttt{url} instead of a \texttt{doi} field: for an \texttt{@article}
without \texttt{volume}/\texttt{pages}, \texttt{apsrev4-x} prints the DOI
as visible text instead of using it as a link target.

\begin{Verbatim}
@article{WilhelmArticleDoi,
    Author = {Wilhelm, Frank K. and ...},
    Title = {An introduction into optimal control for quantum technologies},
    Journal = {arXiv:2003.10132},
    Doi = {10.48550/arxiv.2003.10132},
    Year = {2020},
}
\end{Verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preprint-only entry (HAL) and the \texttt{archive} field}

REVTeX's eprint machinery builds the link text from
\texttt{archivePrefix} and \texttt{eprint}, but its link \emph{target} as
\texttt{<archive>/<eprint>}, where the \texttt{archive} field defaults to
arXiv's \texttt{https://arxiv.org/abs}. For a non-arXiv preprint, the
default is wrong; \texttt{bibdeskparser} exports therefore emit the
\texttt{archive} field with the correct link base (derived from the
archive's URL template in the \texttt{preprint\_archives} configuration).

\subsection{H1: The minimal \texttt{--preprint unpublished} export (the
default)~\cite{TuriniciUnpublished}}

\begin{Verbatim}
@unpublished{TuriniciUnpublished,
    Author = {Turinici, Gabriel},
    Title = {Quantum control},
    Eprint = {hal-00640217},
    Archiveprefix = {HAL},
    Archive = {https://hal.science},
    Url = {https://hal.science/hal-00640217},
    Note = {published in the Encyclopedia of Applied and
            Computational Mathematics, Springer, 2015,
            \url{https://doi.org/10.1007/978-3-540-70529-1_269}},
    Year = {2012},
}
\end{Verbatim}

\subsection{H2: The minimal \texttt{--preprint misc}
export~\cite{TuriniciMisc}}

\begin{Verbatim}
@misc{TuriniciMisc,
    ... same fields as H1, as @misc ...
}
\end{Verbatim}

\subsection{H3: The minimal \texttt{--preprint article}
export~\cite{TuriniciArticle}}

\begin{Verbatim}
@article{TuriniciArticle,
    Author = {Turinici, Gabriel},
    Title = {Quantum control},
    Journal = {HAL:hal-00640217},
    Url = {https://hal.science/hal-00640217},
    Note = {published in the Encyclopedia of Applied and
            Computational Mathematics, Springer, 2015,
            \url{https://doi.org/10.1007/978-3-540-70529-1_269}},
    Year = {2012},
}
\end{Verbatim}

\subsection{H4: Anti-pattern -- as H1, but without the \texttt{archive}
field~\cite{TuriniciNoArchive}}

What every eprint-carrying form renders like for a non-arXiv archive
\emph{without} the \texttt{archive} field: the link points at
\texttt{arxiv.org/abs/hal-00640217}.

\begin{Verbatim}
@unpublished{TuriniciNoArchive,
    ... same fields as H1, without Archive ...
}
\end{Verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Published paper with a preprint}

\subsection{J1: Journal reference with the structured eprint
fields~\cite{KochFull}}

How \texttt{bibdeskparser} stores a published paper with a preprint, and
what both full and minimal exports write.

\begin{Verbatim}
@article{KochFull,
    Author = {Koch, Christiane P. and ...},
    Title = {Quantum optimal control in quantum technologies. ...},
    Journal = epjqt,
    Eprint = {2205.12110},
    Archiveprefix = {arXiv},
    Doi = {10.1140/epjqt/s40507-022-00138-x},
    Pages = {19},
    Volume = {9},
    Year = {2022},
}
\end{Verbatim}

\subsection{J2: The same entry without the eprint
fields~\cite{KochNoEprint}}

What the bibliography loses by dropping \texttt{eprint}/%
\texttt{archiveprefix}: the free-copy link for readers behind a paywall.

\begin{Verbatim}
@article{KochNoEprint,
    ... as J1, without Eprint/Archiveprefix ...
}
\end{Verbatim}

\subsection{J3: A published paper with a HAL preprint, as
exported~\cite{SauvagePRXQ}}

Full and minimal exports of published entries also append the
\texttt{archive} link base when the \texttt{eprint} names a non-arXiv
archive.

\begin{Verbatim}
@article{SauvagePRXQ,
    Author = {Sauvage, Fr{\'e}d{\'e}ric and Mintert, Florian},
    Title = {Optimal Quantum Control with Poor Statistics},
    Journal = {PRX Quantum},
    Year = {2020},
    Doi = {10.1103/prxquantum.1.020322},
    Pages = {020322},
    Volume = {1},
    Number = {2},
    Eprint = {hal-03612955},
    Archiveprefix = {HAL},
    Archive = {https://hal.science},
}
\end{Verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observations}

In the table below, $\langle\ldots\rangle$ marks the extent of a hyperlink
in the rendered entry (follow the links in the bibliography itself to
verify), and ``id'' abbreviates the identifier of the respective
reference. Authors and title precede each rendering.

\begin{center}
\begin{tabular}{lp{0.62\textwidth}}
\hline
P1~\cite{WilhelmStored} &
  $\langle$arXiv:id [quant-ph]$\rangle_{\rm arXiv}$ (2020), lecture
  notes --- identifier in the venue position, category tag, status \\
P2~\cite{WilhelmUnpublished} &
  identical to P1 (the pseudo-journal is ignored) \\
P3~\cite{WilhelmMisc} &
  $\langle$title$\rangle_{\rm DOI}$ (2020), lecture notes,
  $\langle$arXiv:id [quant-ph]$\rangle_{\rm arXiv}$
  --- DOI-linked title; eprint trails after the year \\
P4~\cite{WilhelmArticle} &
  $\langle$arXiv:id (2020)$\rangle_{\rm DOI}$, lecture notes
  --- venue-position identifier under \emph{every} style \\
P5~\cite{WilhelmArxivExport} &
  $\langle$title$\rangle_{\rm arXiv}$ (2020),
  $\langle$arXiv:id [quant-ph]$\rangle_{\rm arXiv}$
  --- like P3, but no DOI and no status \\
P6~\cite{WilhelmLegacyArticle} &
  arXiv:id $\langle$10.48550/arxiv.id$\rangle_{\rm DOI}$ (2020),
  $\langle$arXiv:id [quant-ph]$\rangle_{\rm arXiv}$
  --- identifier shown \emph{three} times, raw DOI visible \\
P7~\cite{WilhelmArticleDoi} &
  arXiv:id $\langle$10.48550/arxiv.id$\rangle_{\rm DOI}$ (2020)
  --- raw DOI visible next to the pseudo-journal \\
\hline
H1~\cite{TuriniciUnpublished} &
  $\langle$HAL:hal-00640217$\rangle_{\rm HAL}$ (2012),
  $\langle$Note$\rangle_{\rm DOI}$ --- eprint link correct,
  thanks to \texttt{archive} \\
H2~\cite{TuriniciMisc} &
  $\langle$title$\rangle_{\rm HAL}$ (2012), $\langle$Note$\rangle_{\rm DOI}$,
  $\langle$HAL:hal-00640217$\rangle_{\rm HAL}$
  --- all links correct \\
H3~\cite{TuriniciArticle} &
  $\langle$HAL:hal-00640217 (2012)$\rangle_{\rm HAL}$,
  $\langle$Note$\rangle_{\rm DOI}$ --- both links correct \\
H4~\cite{TuriniciNoArchive} &
  $\langle$HAL:hal-00640217$\rangle_{\rm broken}$ (2012),
  $\langle$Note$\rangle_{\rm DOI}$ --- without
  \texttt{archive}, the eprint link points at \texttt{arxiv.org} \\
\hline
J1~\cite{KochFull} &
  $\langle$EPJ Quantum Technol.\ \textbf{9}, 19
  (2022)$\rangle_{\rm DOI}$, $\langle$arXiv:id$\rangle_{\rm arXiv}$
  --- journal reference \emph{and} free-copy link \\
J2~\cite{KochNoEprint} &
  as J1, but \emph{without} the arXiv link \\
J3~\cite{SauvagePRXQ} &
  $\langle$PRX Quantum \textbf{1}, 020322 (2020)$\rangle_{\rm DOI}$,
  $\langle$HAL:hal-03612955$\rangle_{\rm HAL}$
  --- the free-copy link is correct, thanks to \texttt{archive} \\
\hline
\end{tabular}
\end{center}

Key observations, specific to \texttt{apsrev4-2} (2019-01-14) with
\texttt{hyperref}:

\begin{enumerate}
\item The stored \texttt{@unpublished} form (P1) can be copied verbatim
  into a document: it renders exactly like the minimal export (P2),
  because BibTeX ignores the \texttt{journal} field of an
  \texttt{@unpublished} entry. The eprint renders hyperlinked in the
  venue position, with the category tag and the status note -- the best
  REVTeX rendering of a preprint. The legacy \texttt{@article} storage
  form (P6) can \emph{not} be copied verbatim; \texttt{bibdeskparser
  export} normalizes either stored form.
\item \texttt{note} is a \emph{required} field of \texttt{@unpublished}:
  BibTeX warns ``empty note'' for every entry without one. Recording the
  publication status there (``preprint only'', ``submitted to Phys.\
  Rev.\ A'', ``lecture notes'') is therefore doubly useful; minimal
  exports synthesize ``preprint'' when an entry stores none.
\item The three export forms are complementary. \texttt{unpublished}
  (P1/P2, H1) and \texttt{misc} (P3, H2) rely on the style rendering the
  \texttt{eprint} field --- which REVTeX, \texttt{elsarticle}, and
  biblatex do, but \texttt{plain}, \texttt{unsrt}, \texttt{abbrv}, and
  \texttt{IEEEtran} do \emph{not} (under those, only the note text
  survives, and the identifier is lost). \texttt{article} (P4, H3)
  renders the identifier in the journal position under \emph{every}
  style, at the cost of the native eprint formatting (no category tag).
\item For an \texttt{@article} without \texttt{volume}/\texttt{pages},
  the \texttt{doi} field is printed as \emph{visible text} (P6, P7)
  rather than becoming the link target of the journal reference, as it
  does when \texttt{volume}/\texttt{pages} are present (J1, J3). The
  \texttt{article} export form therefore writes the DOI as a
  \texttt{url}, which wraps the journal name and year in the hyperlink
  (P4, H3).
\item The link target of REVTeX's rendered eprint is
  \texttt{<archive>/<eprint>}, with the \texttt{archive} field
  defaulting to \texttt{https://arxiv.org/abs}. For a non-arXiv archive
  that default is a broken link (H4); emitting the correct base in
  \texttt{archive} fixes the eprint link everywhere it appears (H1, H2,
  J3). The \texttt{url} field cannot compensate on
  \texttt{@unpublished}, where it is ignored.
\item Keeping the structured \texttt{eprint} fields on a
  \emph{published} entry costs nothing and adds the hyperlinked
  identifier after the journal reference (J1/J3 vs.\ J2) --- the link
  that stays accessible to readers without a journal subscription.
  \texttt{bibdeskparser}'s minimal export therefore includes them
  (and every export adds the \texttt{archive} base where needed, J3).
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{refs}

\end{document}
